---
title: "p8105_hw2_tj2519"
output: github_document
date: "2023-09-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
```


# Problem 1 - process the pols-month, snp, unemployment dataset
## processing the datasets
```{r}
pols_df = 
  read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day"), "-", convert = TRUE) %>% 
  mutate(month = ifelse(month == 1, "jan", month),
         month = ifelse(month == 2, "feb", month),
         month = ifelse(month == 3, "mar", month),
         month = ifelse(month == 4, "apr", month),
         month = ifelse(month == 5, "may", month),
         month = ifelse(month == 6, "jun", month),
         month = ifelse(month == 7, "jul", month),
         month = ifelse(month == 8, "aug", month),
         month = ifelse(month == 9, "sep", month),
         month = ifelse(month == 10, "oct", month),
         month = ifelse(month == 11, "nov", month),
         month = ifelse(month == 12, "dec", month)
         ) %>% 
  mutate(president = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  select(-day, -starts_with("prez"))

print(sapply(pols_df, class))
view(pols_df)



snp_df = 
  read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% 
  separate(date, c("year", "month", "day"), "-") %>% 
  mutate(year = as.numeric(year), month = as.numeric(month)) %>% 
  mutate(year = ifelse(year > as.integer(format(Sys.Date(), "%Y")), 
                       year - 100, year)) %>% 
  arrange(year, month) %>% 
  mutate(month = ifelse(month == 1, "jan", month),
         month = ifelse(month == 2, "feb", month),
         month = ifelse(month == 3, "mar", month),
         month = ifelse(month == 4, "apr", month),
         month = ifelse(month == 5, "may", month),
         month = ifelse(month == 6, "jun", month),
         month = ifelse(month == 7, "jul", month),
         month = ifelse(month == 8, "aug", month),
         month = ifelse(month == 9, "sep", month),
         month = ifelse(month == 10, "oct", month),
         month = ifelse(month == 11, "nov", month),
         month = ifelse(month == 12, "dec", month)
         ) %>%
  select(-day)
  
print(sapply(snp_df, class))
view(snp_df)


unemployment_df = 
  read_csv("data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percent"
  )

print(sapply(unemployment_df, class))  
view(unemployment_df)



```

## left join the datasets
```{r}
data_join = left_join(pols_df, snp_df) %>% 
  left_join(., unemployment_df)

view(data_join)
```



# Problem 2 
## part1 - read, clean and create the new weight_tons variable in the Mr. Trash Wheel sheet
```{r}
trawh_df = 
  readxl::read_excel("data/202309 Trash Wheel Collection Data.xlsx", 
                     sheet = "Mr. Trash Wheel", range = "A2:N586") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(homes_powered = weight_tons * 500) %>% 
  mutate(wheel_type = "mr_trashwheel") %>% 
  relocate(wheel_type)
  
view(trawh_df)
```

## part2 -  read, clean and create the new weight_tons variable in the Professor Trash Wheel sheet
```{r}
prof_trawh_df = 
  readxl::read_excel("data/202309 Trash Wheel Collection Data.xlsx", 
                     sheet = "Professor Trash Wheel", range = "A2:M109") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(homes_powered = weight_tons * 500) %>% 
  mutate(wheel_type = "prof_trashwheel") %>% 
  relocate(wheel_type)

 view(prof_trawh_df) 
```

## part3 -  read, clean and create the new weight_tons variable in the  Gwynndal sheet
```{r}
gwy_trawh_df = 
  readxl::read_excel("data/202309 Trash Wheel Collection Data.xlsx", 
                     sheet = "Gwynnda Trash Wheel", range = "A2:L159") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(homes_powered = weight_tons * 500) %>% 
  mutate(wheel_type = "gwy_trashwheel") %>% 
  relocate(wheel_type)

 view(gwy_trawh_df) 
```

## part 4 - combine the datasets
```{r}

```

